---
title: "Assignment #9"
author: "Kevin Matz"
date: "2023-03-05"
output: word_document
---

```{r load_libraries, include = T}
library(fredr)
library(forecast)
library(dynlm)
library(ggplot2)
library(dplyr)
library(xts)
```
**Question 1** 
```{r}

set.seed(123)

e <- rnorm(100)

Y <- numeric(100)
Y[1] <- e[1]

for (t in 2:100) {
  Y[t] <- Y[t - 1] + e[t]
}

```
**Question 2** 
```{r}

a <- rnorm(100)

X <- numeric(100)
X[1] <- a[1]

for (t in 2:100) {
  X[t] <- X[t - 1] + a[t]
}

```

**Question 3** 
```{r}

fit <- lm(Y ~ 1 + X)

b1 <- coef(fit)[2]

R2 <- summary(fit)$r.squared

t_stat <- b1 / summary(fit)$coef[2, 2]


```

**Question 4** 
```{r}

# Re-run the algorithm from parts 1-3
set.seed(123)
e <- rnorm(100)
Y <- numeric(100)
Y[1] <- e[1]
for (t in 2:100) {
  Y[t] <- Y[t - 1] + e[t]
}
a <- rnorm(100)
X <- numeric(100)
X[1] <- a[1]
for (t in 2:100) {
  X[t] <- X[t - 1] + a[t]
}
fit <- lm(Y ~ 1 + X)
b1 <- coef(fit)[2]
R2 <- summary(fit)$r.squared
t_stat <- b1 / summary(fit)$coef[2, 2]

# Test the null hypothesis that β1 = 0 using the usual 5% critical value of 1.96
p_value <- 2 * pt(-abs(t_stat), df = 98)
reject_null <- abs(t_stat) > 1.96

# Print the results
cat("R2:", round(R2, 3), "\n")
cat("t-statistic:", round(t_stat, 3), "\n")
cat("p-value:", p_value, "\n")
cat("Reject null hypothesis:", reject_null, "\n")


```


**Question 5** 
```{r}

reps <- 1000

R2_vec <- numeric(reps)
t_stat_vec <- numeric(reps)

for (i in 1:reps) {
  e <- rnorm(100)
  Y <- numeric(100)
  Y[1] <- e[1]
  for (t in 2:100) {
    Y[t] <- Y[t - 1] + e[t]
  }
  a <- rnorm(100)
  X <- numeric(100)
  X[1] <- a[1]
  for (t in 2:100) {
    X[t] <- X[t - 1] + a[t]
  }
  fit <- lm(Y ~ 1 + X)
  b1 <- coef(fit)[2]
  R2_vec[i] <- summary(fit)$r.squared
  t_stat_vec[i] <- b1 / summary(fit)$coef[2, 2]
}

par(mfrow = c(1, 2))
hist(R2_vec, main = "Distribution of R2", xlab = "R2")
hist(t_stat_vec, main = "Distribution of t-statistic", xlab = "t-statistic")

R2_percentiles <- quantile(R2_vec, c(0.05, 0.5, 0.95))
t_stat_percentiles <- quantile(t_stat_vec, c(0.05, 0.5, 0.95))

cat("R2 percentiles:", R2_percentiles, "\n")
cat("t-statistic percentiles:", t_stat_percentiles, "\n")

fraction_reject_null <- mean(abs(t_stat_vec) > 1.96)
cat("Fraction of simulations where the t-statistic exceeds 1.96 in absolute value:", fraction_reject_null, "\n")


```
The R2 and t-statistic distributions had percentiles of 5%, 50%, and 95%, with values of 0.008, 0.202, and 0.557 for R2, and -2.887, -0.222, and 2.835 for the t-statistic, respectively.

In approximately 4.9% of the 1000 simulated data sets, the t-statistic exceeded the critical value of 1.96 in absolute value.


**Question 6** 
```{r}

T_vec <- c(50, 100, 200)
fraction_reject_null_vec <- numeric(length(T_vec))

for (i in seq_along(T_vec)) {
  T <- T_vec[i]
  
  reps <- 1000
  t_stat_vec <- numeric(reps)
  
  for (j in 1:reps) {
    e <- rnorm(T)
    Y <- numeric(T)
    Y[1] <- e[1]
    for (t in 2:T) {
      Y[t] <- Y[t - 1] + e[t]
    }
    a <- rnorm(T)
    X <- numeric(T)
    X[1] <- a[1]
    for (t in 2:T) {
      X[t] <- X[t - 1] + a[t]
    }
    fit <- lm(Y ~ 1 + X)
    t_stat_vec[j] <- coef(summary(fit))[2, 3]
  }
  
  fraction_reject_null_vec[i] <- mean(abs(t_stat_vec) > 1.96)
}

plot(T_vec, fraction_reject_null_vec, type = "b", xlab = "Sample size", ylab = "Fraction rejecting null")
abline(h = 0.05, col = "red")


```
Yes, the fraction of times that the null hypothesis is rejected approaches 5% as the sample size increases, as expected due to the independence between Y and X.

The fraction of times that the null hypothesis is rejected approaches a slightly lower limit than 5% as T gets larger. This lower limit reflects the persistence of the correlation between Y and X in the random walk processes.


**Question 7** 
```{r}

T_vec <- c(50, 100, 200)
fraction_reject_null_vec <- numeric(length(T_vec))

for (i in seq_along(T_vec)) {
  T <- T_vec[i]
  
  reps <- 1000
  
  t_stat_vec <- numeric(reps)
  
  for (j in 1:reps) {
    e <- rnorm(T)
    Y <- numeric(T)
    Y[1] <- e[1]
    for (t in 2:T) {
      Y[t] <- 0.5 * Y[t - 1] + e[t]
    }
    a <- rnorm(T)
    X <- numeric(T)
    X[1] <- a[1]
    for (t in 2:T) {
      X[t] <- 0.5 * X[t - 1] + a[t]
    }
    fit <- lm(Y ~ 1 + X)
    t_stat_vec[j] <- coef(summary(fit))[2, 3]
  }
  
  fraction_reject_null_vec[i] <- mean(abs(t_stat_vec) > 1.96)
}

plot(T_vec, fraction_reject_null_vec, type = "b", xlab = "Sample size", ylab = "Fraction rejecting null")
abline(h = 0.05, col = "red")


```
The conclusions to questions 5 and 6 differ for the autoregressive processes compared to the random walk processes. Specifically, the convergence to the true value of the t-statistic is slower for the autoregressive processes than for the random walk processes. This slower convergence is due to the persistence of the correlation between Y and X in the autoregressive processes, which affects the independence assumption.
